---
title: "part-1"
author: "Dan Turner"
date: "7/28/2021"
output: html_document
---

# Part 1: Cleanining, extracting, tokenization

## Install packages if needed
```{r inistallatiion, message=FALSE, warning=FALSE}
# install.packages("stringi")
# install.packages("stringr")
# install.packages("gutenbergr")
# install.packages("tidytext")
# install.packages("rvest")
# install.packages("topicmodels")
# install.packages("text2vec")
# install.packages("textdata")
# install.packages("stm")
# install.packages("quanteda")
# install.packages("quanteda.textplots")
# install.packages("quanteda.textstats")
# install.packages("purrr")
```

## Setup R for Part 1

```{r}
# Load packages for part 1
library(dplyr) # modern syntax
# library(tokenizers) # consistent, convenient text tokenization
library(gutenbergr) # easy access to gutenberg digital library
library(stringr) # most common string processing functions (uses stringi)
library(tidytext) # most common text processing functions
library(rvest)  # web scraping
library(textdata)


# Load packages for part 3
# library(topicmodels)  # topic modeling
# library(quanteda) # new and powerful text analysis package for R
# library(text2vec)  # map ngram models in vector space to train CNNs
# library(quanteda.textstats)
# library(quanteda.textplots)

```

## Basics of working with strings/text in R

```{r}

# Find the gutenberg_id of a book to scrape
# Return only books where 'Doyle' is detected in the 'author' position
gutenberg_ids <- gutenberg_works(str_detect(author, "Doyle"))

# For example
gutenberg_sherlock <- gutenberg_download(108)$text # run this and check the result

# Download the raw text of your choice
# gutenberg_text <- gutenberg_download(108)$text

# A non-generalized way of doing the same thing
# gutenberg_ids <- gutenberg_works(author == "Doyle, Arthur Conan")

# Here are some ids of popular books:
# 1342 = Pride and Prejudice by Jane Austen
# 11 = Alice's Adventures in Wonderland by Lewis Carroll
# 64317 = The Great Gatsby by F. Scott Fitzgerald

```

## Checking the source

How clean is the data source? Do you consider the text before the first sentence of a book as part of the book, or do you want to restrict yourself to paragraphs of text? The first step is seeing how the data is formatted.

```{r sherlock}

# For The Return of Sherlock Holmes, the text begins on line 33
head(gutenberg_download(108)$text, 40)

```

## Extracting the 'adventures'

```{r}

# Old fashioned way

# The lines we want
gutenberg_sherlock[16:28] # note leading whitespace

# Import adventure list as a clean tibble
sherlock_adventures <- gutenberg_sherlock[16:28] %>%
  trimws() %>% # trim extra whitespace around each line 
  str_replace_all("[\\p{P}\\p{S}&&[^-]]", "")# replace all punctuation (p) with a space (second argument) except (Subtract) hyphens


# Compare
gutenberg_sherlock[21] # raw
sherlock_adventures[6] # clean


# We could also find these using a regular expression (regex) -- more on that later
```

Next steps: Extract the text of The Adventure of the Missing Three-Quarter.

```{r missing-three-quarter}

# Convert the text to lowercase
sherlock_adventures <- tolower(sherlock_adventures)
gutenberg_sherlock <- tolower(gutenberg_sherlock)

# Find the part of the file to extract (end of target title to beginning of following title)
target_title <- sherlock_adventures[11] # The Adventure of the Missing Three-Quarter is the 11th element in the table of contents we scraped

following_title <- sherlock_adventures[12] # The title of the next story will help us know where to stop

# Now let's find the boundaries of the text span we want to extract

# *Part 1* Get the start of the span (end of the title)

# Detect the target title in the text
hits <- str_detect(gutenberg_sherlock, target_title)

# Which "hit" do we want to extract?
gutenberg_sherlock[hits] # 2 instances, 1 in table of contents, 2 in the text

# Use str_locate_all to locate every match we detected
hits <- as.data.frame(cbind(1:length(hits),
                               hits,
                               str_locate_all(gutenberg_sherlock,
                                              target_title))) 

colnames(hits) <- c("line", "detected", "location")

start <- hits %>% filter(detected == TRUE) %>%
  select(line) %>% # select the line number col
  last() %>% # to access the line numbers inside the col
  nth(2) # second element from gutenberg_sherlock[detected]

# *Part 2* Get the end of the span (following title)
hits <- str_detect(gutenberg_sherlock, following_title)

# Which "hit" do we want to extract?
gutenberg_sherlock[hits] # 3 instances, 1 in table of contents, 2 in the text, 3 is something else

# Wrap the hit and span info in a dataframe again as we locate the hits
hits <- as.data.frame(cbind(1:length(hits),
                               hits,
                               str_locate_all(gutenberg_sherlock,
                                              following_title))) 

colnames(hits) <- c("line", "detected", "location")

end <- hits %>% filter(detected == TRUE) %>%
  select(line) %>% # select the line number col
  last() %>% # to access the line numbers inside the col
  nth(2) # second element from gutenberg_sherlock[detected]

# Extraction
sherlock_raw <- gutenberg_sherlock[start + 1 : end - 1]

head(sherlock_raw) # Peek

# What additional processing do you think we need?

# Make a tokenized version for analysis later (as a tibble, too use tidytext)
sherlock_tokens <- tibble(txt = sherlock_raw[3:length(sherlock_raw)]) %>% # trim title
  unnest_tokens(word, txt) # tidytext tokenization by word (space)

head(sherlock_tokens)

```


# HTML Tables: Scrape & clean a Wikipedia page table

Data displayed in HTML tables can be easily converted to date frame format.

```{r scrape wikipedia table}

# Timeline of geopolitical changes (1900−present)
wiki_url <- "https://en.wikipedia.org/wiki/Timeline_of_geopolitical_changes_(1900%E2%88%92present)"

wiki_html <- read_html(wiki_url)

# Extract the tables from the HTML and combine them as a dataframe
wiki_tables <- wiki_html %>%  # read the html
 html_nodes("#mw-content-text > div.mw-parser-output > table") %>%  # css selector rule
 html_table() %>%  # convert HTML table to a df
 bind_rows(.id = "tbl_order") # combine all the dfs, add new col

# Check result
wiki_tables %>% sample_n(10)

# Simplify the dataset
wiki_tables <- wiki_tables %>%
  select(c("Year", "Event")) %>% # Keep only these columns
  na.omit() # drop the leading blank row

# Check result
wiki_tables %>% sample_n(10)

wiki_tokens <- wiki_tables %>%
  group_by(Year) %>%
  summarize(Event_By_Year = paste0(Event, collapse = ""))  %>% # one row per year
  unnest_tokens(word, Event_By_Year) # tokenize while retaining year

# Check result
wiki_tokens %>% sample_n(10)
```

# Quick sentiment analysis

```{r}

# Approve download of this dataset if given the option
sentiments <- get_sentiments("afinn")

# Join the rated sentiments with the tokens and summarize by year (optional)
wiki_sentiment <- wiki_tokens %>%
  group_by(Year) %>%
  inner_join(sentiments) # %>% summarise(sentiment = mean(value))

sum(wiki_tokens$)

```

# Review

We have acquired some different kinds of text from books and Wikipedia, did some data preparation, and some data extraction.

# What we didn't cover on the topic of cleaning
Regular expressions (regex) are a powerful tool for text processing.
Stopword removal will be covered in Part 3.

## Save progress
```{r}

# You'll load these files in Part 3.
saveRDS(sherlock_raw, "part-1-sherlock-raw.rds")
saveRDS(sherlock_tokens, "part-1-sherlock.rds")
saveRDS(wiki_tokens, "part-1-wiki.rds")

```



